\begin{abstract}
The lack of proprioception makes it challenging for users to use a real keyboard in virtual reality.
This paper proposes SwipeVR, a virtual keyboard that requires only a trackpad, which are available in existing VR input devices today.  The key design principles underlying SwipeVR are: (1) Relative positioning: the user's finger is pointing to the same position in a keyboard upon first contact.
(2) Keybatching with smart correction: the keyboard is partitioned into 8 tiles.  The user does not have to hit the right key, but just the tile that contains the key.
(3) Intention-based mixed keyboard.  The user can also type on the full keyboard - the intent in the choice of the batched vs full keyboard is automatically inferred.
This paper performs a user study across the various design choices as well as existing input methods.
We found that SwipeVR is easy to learn and provides a similar speed to text entry on a mobile phone keyboard.



We present what we believe is the most efficient and fastest text entry method available for virtual reality.
SwipeVR strives for simplicity while maintaining good performance.
For full words, the user gestures on the touch pad of the controller wand to a sector of the keyboard.
Full access to characters, punctuation, and digits are accessible through a secondary keyboard.
Error correction and editing are supported with multi-wand gestures.
The system is able to be learned quickly by beginners and supports eyes-free entry by experts.
The effort to enter text is comparable to gesture text entry on a mobile phone.
We conduct a study with X participants where an average text entry speed of Y WPM was observed with an error rate of Z\%.  
These results show that text entry is similar to that of a mobile phone keyboard but error correction and editing, the most time consuming step, is greatly improved.




\end{abstract}

\category{H.5.1.}
{Information Interfaces and Presentation (e.g. HCI)}
{Multimedia Information Systems}
\category{I.3.7.}
{Computer Graphics}
{Three-Dimensional Graphics and Realism}

%link at: http://www.acm.org/about/class/ccs98-html

\keywords{\plainkeywords}


